### 2018-11-04 paper/complexity.tex
Use normalised transition graphs to explain how Z is modified with internal choice STOPs

% -----------------------------------------------------------------------------
\subsubsection*{Test Cases of $\pmax$}
The test cases $U_F(j)$ generated from $\pmax$ 
can never enter branch (\ref{eq:ufa}), because $\pmax/s$ has initials $\Sigma$ for
all traces $s\in\trc(\pmax)$ according to Lemma~\ref{lemma:pmax}, (\ref{eq:pmaxa}).
Moreover, they can never enter branch (\ref{eq:ufb}), because $\minhits(\pmax/s)$
is never empty according to Lemma~\ref{lemma:pmax}, (\ref{eq:pmaxd}).
Finally, the minimal hitting sets used to probe the SUT at the end of a
non-blocking  test execution are always the hitting sets of ${\cal C}$ 
according to Lemma~\ref{lemma:pmax}, (\ref{eq:pmaxe}).
This results in a test case structure
\begin{eqnarray*}
\label{eq:UFP}
U_F(j) & = & U_F(j,0,\ii n)
\\
U_F(j,k,n) & = &   (k < j) \& \big(e:\Sigma   \then U_F(j,k+1,t(n,e) \big)
\\ & & \extchoice  
\\ & & (k = j) \& \big( \Intchoice_{H\in\minhits({\cal C})} (e:H   \then \epass \then\Stop   )  \big)
\end{eqnarray*}

@@

It remains to 
consider the question whether reference processes $Z$ possessing failures $(s,\Sigma)$
may require more test executions for their associated tests $U_F(j)$ than the bound
given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider an {\it arbitrary} process  $Z$ and study the potential increase of test executions if
the behaviour of $Z$ is changed to $(Z/\ol s)\intchoice \Stop$ after some specific 
trace $\ol s\in\trc(Z)$ with $\minhits(Z/\ol s) \neq \varnothing$. This modified process is denoted by $Z({\ol s})$
an can be specified by\footnote{We use the following prefix operators
for given   traces $s_1,s_2\in\Sigma^*$: the Boolean 
expression $ s_1 < s_2$ evaluates to $\ist$ if and only if $s_1$ is a true prefix of $s_2$. Expression $s_1\le s_2$ holds if and only if $ s_1 < s_2 \vee s_1 = s_2$ is true.}
\begin{eqnarray}
Z({\ol s}) & = & Z({\ol s},\varepsilon)
\\
\label{eq:zstop}
Z(\ol s,s) & = & (s = \ol s) \& ((Z/s)\intchoice \Stop)
\\ & & \Extchoice    \nonumber
\\  
\label{eq:zstoppref}
& & (s < \ol s \wedge \minhits(Z/s) = \varnothing) \& 
\big(  (e:[Z/s]^0 \then  Z(\ol s, s.e)) \intchoice \Stop \big)
\\ & & \Extchoice   \nonumber
\\ 
\label{eq:zcont}
& &           (s < \ol s\wedge \minhits(Z/s) \neq \varnothing ) \& 
\\ & & \nonumber
\big(\Intchoice_{A\in\minaccs(Z/s)\cup \{ [Z/s]^0 \}}
e : A \then Z(\ol s, s.e)          \big)
\\ & & \Extchoice   \nonumber
\\
\label{eq:zlikep}
& & (s \not\le \ol s) \& (Z/s)
\end{eqnarray}
%
If $Z(\ol s,s)$ has executed trace $\ol s$, its branch (\ref{eq:zstop}) is entered, 
so the process 
nondeterministically decides whether to continue like $Z/s$ or to stop. As a consequence,
$Z(\ol s)/\ol s = Z(\ol s,\ol s)$ only has $\varnothing$ as minimal acceptance, and its
  collection of minimal hitting sets is empty.
 While $Z(\ol s)$ still executes a true prefix
of $\ol s$, it behaves exactly as $Z$, executing either branch (\ref{eq:zstoppref}) or
(\ref{eq:zcont}): the latter branch is taken if $Z/s$ has at least one minimal hitting set,
so that it can never refuse all elements of $[Z/s]^0$. In such a case, $Z/s$ can be
represented in a normalised form, where an initial event is selected from an internal
choice over all minimal acceptances and over $[Z/s]^0$ (the choice $[Z/s]^0$ is needed
because the union over all minimal acceptances is not necessarily identical 
to $[Z/s]^0$)~\cite[pp.~277]{Roscoe:1997:TPC:550448}. 
The former branch is taken if $Z/s$ can block all of its initials.
As a consequence, in both branches, $Z(\ol s,s)$ 
has the same nondeterministic behaviour and the same initials as $Z/s$, as long as
a prefix of $\ol s$ is executed. As soon as $Z(\ol s)$ has run through a trace $s$ that is no longer a prefix of $\ol s$, 
the process continues exactly as $Z$ would do after having run through $s$. This situation
is captured by branch (\ref{eq:zlikep}).
Summarising, $Z(\ol s)$ behaves like $P$ for all but the
one trace   $\ol s$.

Let us now  compare test executions of $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ for
various values of $j$, where ${\ol U}_F(j)$ denotes a failures test case derived from 
reference process $Z(\ol s)$, to executions of $(Z\parallel[\Sigma] U_F(j))$, where
$U_F(j)$ has been derived from $Z$.
\begin{enumerate}
\item For $j<\#\ol s$, $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ has the same executions 
as $(Z\parallel[\Sigma] U_F(j))$.

\item For $j = \#\ol s$, $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ has 
$(\card{\minhits(Z/\ol s)} - 1)$ fewer executions than $(Z\parallel[\Sigma] U_F(j))$, because 
$Z(\ol s,\ol s)$ has an empty set of hitting sets, since 
$Z(\ol s,\ol s) = (Z/\ol s)\intchoice\Stop$. As a consequence, test case
${\ol U}_F(\#\ol s)$ can only enter branches 
(\ref{eq:ufa})  or (\ref{eq:ufb}) after having run through trace
$\ol s$. The former branch is available if and only if it is available in $U_F(\#\ol s)$, the latter just stops the test after issuing the $\epass$ event.

\item For $j > \#\ol s$, one execution is won by entering 
branch (\ref{eq:ufb}) after $\ol s$, no executions are lost.
\end{enumerate}
%
Summing up over all test cases for $j=0,\dots,pq-1$, the difference in test executions generated
from $Z(\ol s)$ and those generated from  $Z$ is
\begin{equation}
 pq - \#\ol s       - \card{\minhits(Z/\ol s)}.
\end{equation}
%
This expression is maximal for $\card{\minhits(Z/\ol s)} = 1$ (recall that
the definition of $Z(\ol s)$ requires $\card{\minhits(Z/\ol s)} > 0$). 
The number of hitting sets in $\minhits(Z/\ol s)$ is 1, for example, if   the
minimal acceptances of $Z/\ol s$ have exactly one element in their common intersection.
The maximal length of 
$\ol s$ where the expression remains strictly positive is then $pq - 2$.

The construction of $Z(\ol s)$ can be generalised by using $k$ traces 
${\ol s}_i, \ i = 1\dots,k$ of the same length $\ol j$, 
after that the modified process behaves as 
$(Z/{\ol s}_i)\intchoice \Stop$. Generalising the case distinction above, 
\begin{enumerate}
\item the number of test executions remains unchanged for $j < \ol j$,
\item the number of   executions lost for $j = \ol j$ is at least 
$k$ (assuming that for all $i$, $\card{\minhits(Z/{\ol s}_i)} = 1$), and
\item the number of executions won for $j > \ol j$ is $k$.
\end{enumerate}
%
The number $k$ of traces ${\ol s}_i,\ i=1,\dots,k$ with fixed length $\#{\ol s}_i=\ol j$ is bounded by the maximal number
of traces that can be executed by $(Z\parallel[\Sigma] U_F(\ol j))$, which is bounded by
$\card{\Sigma}^{\ol j}$. If $k$ equals this maximal number, there is no need
to execute test case ${\ol U}_F(\ol j)$ any longer, because it can only stop after $\epass$, whenever a trace of length $\ol j$ is executed.

Summing up over all test cases ${\ol U}_F(j),\ j = \ol j,\dots,pq-1$, for a given 
 $k \le \card{\Sigma}^{\ol j}$, this results in 
an upper bound of
\begin{equation}
 \card{\Sigma}^{\ol j}(pq - \ol j      - 1)
\end{equation}
additional test executions in comparison to the number of executions needed for all tests
of $Z$.

So far, the additional executions originating from traces with fixed length $\ol j\in\{0,\dots,pq-2\}$ have been considered. Summing up over all lengths $\ol j$ results in
an upper bound of 
%
\begin{equation}
\sum_{j=0}^{pq-2} n^j(pq - j - 1) = \frac{n^{pq} - npq + pq - 1}{(n-1)^2}\quad\text{with}\ 
n=\card{\Sigma}
\end{equation}
%
additional test executions. This has the order of magnitude $O(n^{pq-2})$. Therefore, the additional test executions won by being able to enter branches (\ref{eq:ufb}) in test executions  can be neglected in comparison to the order of magnitude calculated for $\pmax$ in (\ref{eq:maxO}).  
These considerations lead to the following theorem
%
\begin{theorem}
Given a process alphabet $\Sigma$ with cardinality $
\card{\Sigma} = n$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
the normalised transition graph of $P$ has $p$ states, and the fault domain 
${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then the maximal 
number of test executions to be performed   using the complete test 
suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ as specified in Theorem~\ref{th:failurestest} is of order
%
\begin{equation*} 
O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot n^{pq-1}\big)\quad\text{with $n=\card{\Sigma}$}.
\end{equation*} 
\xbox
\end{theorem}
%
In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test {\it every}
non-empty subset of $\Sigma$ whose events cannot be completely refused in a
given process state of the reference model; this leads to a worst-case
estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be
offered to the SUT in the last step of the test execution. This number is significantly larger than the worst-case estimate $\binom{n}{\lfloor n/2\rfloor}$ calculated above for the 
hitting sets to be checked.
In Fig.~\ref{fig:minhita}, the reduction is visualised by plots of
the two functions.

>>>>

% -----------------------------------------------------------------------------
\subsubsection*{Test Cases of $\pmax$}
The test cases $U_F(j)$ generated from $\pmax$ 
can never enter branch (\ref{eq:ufa}), because $\pmax/s$ has initials $\Sigma$ for
all traces $s\in\trc(\pmax)$ according to Lemma~\ref{lemma:pmax}, (\ref{eq:pmaxa}).
Moreover, they can never enter branch (\ref{eq:ufb}), because $\minhits(\pmax/s)$
is never empty according to Lemma~\ref{lemma:pmax}, (\ref{eq:pmaxd}).
Finally, the minimal hitting sets used to probe the SUT at the end of a
non-blocking  test execution are always the hitting sets of ${\cal C}$ 
according to Lemma~\ref{lemma:pmax}, (\ref{eq:pmaxe}).
This results in a test case structure
\begin{eqnarray*}
\label{eq:UFPpmax}
U_F(j) & = & U_F(j,0,\ii n)
\\
U_F(j,k,n) & = &   (k < j) \& \big(e:\Sigma   \then U_F(j,k+1,t(n,e) \big)
\\ & & \extchoice  
\\ & & (k = j) \& \big( \Intchoice_{H\in\minhits({\cal C})} (e:H   \then \epass \then\Stop   )  \big)
\end{eqnarray*}

@@

It remains to 
consider the question whether reference processes $Z$ possessing failures $(s,\Sigma)$
may require more test executions for their associated tests $U_F(j)$ than the bound
given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider an {\it arbitrary} process  $Z$ represented by its transition graph $G(Z)$ with $p$ nodes. 
Let $v$ be any node of $G(Z)$, such 
that $\minaccs(v)\neq\{\varnothing \}$. Now modify $G(Z)$ by setting $v$'s acceptance label
to $\{\varnothing \}$. This produces a modified node $\ol v$ with $[\ol v]^0=[v]^0$, but
$\minaccs(\ol v)=\{\varnothing \}$. Now it is checked whether $\ol v$ is equivalent to any other node of $G(Z)$. If this is not the case, the modified transition graph $G(\ol Z)$ constructed by
exchanging $v$ with $\ol v$ and keeping all other nodes and all edges of $G(Z)$  in the new graph without changes has exactly the same number of nodes as $G(Z)$.

Now consider a test case ${\ol U}_F(j)$ for $j\in\{ 0,\dots,pq-1 \}$ derived from $G(\ol Z)$
and compare it to $U_F(j)$ derived from $G(Z)$. Let $s$ be a trace of  ${\ol U}_F(j)$ such that
$\#s\le j$ and   $G(\ol Z)/s = \ol v$. Then we have two cases to distinguish.
\begin{enumerate}
\item If $\#s < j$, then ${\ol U}_F(j)/s$ allows for one additional execution step entering branch 
(\ref{eq:ufb}) of the test, due to construction of $\ol v$.

\item If $\#s = j$, then ${\ol U}_F(j)/s$ loses $\card{\minhits(v)}$ execution steps checking 
that $Z/s$ never blocks one of the hitting sets in $\minhits(v)$ and wins one execution step 
because again, branch 
(\ref{eq:ufb}) of the test can now be entered. The difference $1 - \card{\minhits(v)}$ has 0 as
the largest possible value. This applies, for example, if the intersection of 
all minimal acceptances of $n$ contains exactly one element.
\end{enumerate}

As a very rough estimate for the maximal number of additional test executions of ${\ol U}_F(j)/s$
in comparison to $U_F(j)$ we can therefore assume that $\card{\minhits(v)} = 1$ and all
$\card{\Sigma}^k$ traces of length $k < j$ lead to the new node $\ol v$. As a result,
${\ol U}_F(j)/s$ has one additional execution step for every trace of length less than $j$.  
This results in an upper bound of 
\begin{equation}
\sum_{i=0}^{j-1}n^i = \frac{n^j - 1}{n - 1} \quad\text{with}\ n = \card{\Sigma}
\end{equation}
test executions. Summing up over additional test executions stemming from al test cases 
 ${\ol U}_F(j),\ j = 0,\dots,pq-1$ results in the following upper bound for the number of 
 additional executions.
\begin{equation}
\sum_{j=0}^{pq-1}\frac{n^j - 1}{n - 1} = \frac{n^{pq} - npq + pq - 1}{(n-1)^2}\quad\text{with}\ 
n=\card{\Sigma}.
\end{equation}
This expression has the order of magnitude $O(n^{pq-2})$.  Therefore, the additional test executions won by being able to enter branches (\ref{eq:ufb}) in test executions  can be neglected in comparison to the order of magnitude calculated for $\pmax$ in (\ref{eq:maxO}).

These considerations lead to the following theorem
%
\begin{theorem}
Given a process alphabet $\Sigma$ with cardinality $
\card{\Sigma} = n$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
the normalised transition graph of $P$ has $p$ states, and the fault domain 
${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then the maximal 
number of test executions to be performed   using the complete test 
suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ as specified in Theorem~\ref{th:failurestest} is of order
%
\begin{equation*} 
O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot n^{pq-1}\big)\quad\text{with $n=\card{\Sigma}$}.
\end{equation*} 
\xbox
\end{theorem}
%
In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test {\it every}
non-empty subset of $\Sigma$ whose events cannot be completely refused in a
given process state of the reference model; this leads to a worst-case
estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be
offered to the SUT in the last step of the test execution. This number is significantly larger than the worst-case estimate $\binom{n}{\lfloor n/2\rfloor}$ calculated above for the 
hitting sets to be checked.
In Fig.~\ref{fig:minhita}, the reduction is visualised by plots of
the two functions.
